Hydrogen Epoch of Reionization Array (HERA)

### 2016-04-25 manuscript.tex
Finish reading through for typos.

The past three years have seen deep EOR observations with PAPER, the MWA, 
the Giant Metrewave Radio Telescope (GMRT), and LOFAR.
PAPER and the MWA have produced progressively deeper limits
\citep{dillon_et_al2014,dillon_et_al2015,parsons_etal2014,ali_et_al2015}, with PAPER
yielding the first meaningful constraints on the 21\,cm spin temperature during reionization
(Fig. \ref{fig:limits}). %XXX: This figure doesn't really say much about spin temperature.
The inherent challenge is to simultaneously meet stringent sensitivity requirements 
while suppressing foregrounds $\sim5$ orders of magnitude
brighter than the 21\,cm signal, which has been addressed along a number of approaches.  
To address this challenge,  all aspects of the experimental process must be refined and improved, spanning 
calibration ({\em e.g.} \cite{zheng_et_al2014, jacobs_et_al2016, 2016arXiv160300607B}), handling foreground contamination
({\em e.g.} \cite{moore_et_al2013,thyagarajan_et_al2015a,pober_et_al2016}),
and the interferometer design itself ({\em e.g.} \cite{parsons_et_al2012a,dillon_parsons2016}).

@@

The $\kvpr$ components are directly proportional to the baselines and $\kpar$ are proportional to the Fourier transform of the frequency response \footnote{Note that this equivalency is an approximation, which is very good for the short baselines and small bandwidths generally used here.  For a more complete discussion see \cite{liu_et_al2014a,liu_et_al2014b}}.   
The Fourier transform of a frequency spectrum is a delay, hence the name ``delay-spectrum'' for the technique that uses this approach.  Note that cosmic evolution limits the largest bandwidth (which determines the smallest $\kpar$) to about 10 MHz -- for larger bandwidths the evolution of the Universe begins to impact the result.  For HERA, wavenumbers are dominated by the bandwidth, not the baseline.

Figure \ref{fig:kperf} shows many of the dependencies on wavenumber by the bandwidth, configuration and cosmology by plotting the perpendicular wavenumber ($\kvpr$; black lines), parallel wavenumber ($\kpar$; blue lines) and total wavenumber ($k$; green lines) as a function of redshift for various bandwidths and baselines.  The redshift range is appropriate for the extended frequency range that is the goal for HERA, however $z$=6-13 is the primary region for EOR studies.
The lines are labelled for the assumed bandwidth and/or baseline, depending on the type of wavenumber shown.  Obviously many values may be used, but here a small sampling of values appropriate for HERA are shown.    
Wavenumbers with an assumed bandwidth ($k$ and $\kpar$ are shown as stepped profiles, with the redshift steps equal to the bandwidth at that redshift.
Note that for 100 MHz, the concept of a power spectrum at a given redshift is no longer appropriate, which is why the line is dashed.
Also in Fig. \ref{fig:kperf}, the lime-green shaded area indicates the initial region in which HERA intends to detect and characterize the EOR using the delay-spectrum technique.

@@

If this $\beta$ factor is determined by the chromaticity of the longest baseline, we see that $\beta=1$ and this line is referred to as the ``horizon line'' ({\em i.e.}, the delay of a source at the horizon).  Systematics will push that line such as to close the EOR ``window'' to the upper left ($\beta > 1$).  If we can completely control systematics and constrain the affects of the foreground to {\em e.g.} the main beam, the line will move to decrease the size of the wedge ($\beta<1$).
Therefore, a key issue to measuring the EOR power spectrum is to understand and minimize systematics.  For further discussion and an analytical derivation, see \cite{zahn_etal2012,vedantham_2012,liu_et_al2014b},  in simulation see \cite{datta_etal2010,hazelton_et_al2013} and for observations
\cite{pober_etal2013b,2015arXiv150601026P,parsons_etal2014,2015arXiv150206016A}.

Note that the unit of the spatial wavenumber $k$ is 1/length, in this case the relevant length scale is megaparsecs (Mpc; 1 Mpc = 3.086$\times10^{22}$ m).  In order to normalize for updated measurements of the Hubble constant, it is further normalized by a factor $h$, where Hubble's constant is expressed as $H_0=100h$ (km/s)/Mpc , such that the wavenumbers are expressed as $h$/Mpc.  Also note that at the redshifts of interest $X$ has a value of about 160 Mpc/deg and $Y$ has a value of about 16 Mpc/MHz.

@@

The techniques to measure the power spectrum may be broken down into two principle techniques -- delay-space (\S\ref{sec:delayapproach}) and map-making (\S\ref{sec:mapapproach}) -- plus additional hydrid methods.  These are briefly summarized below, with an emphasis on the delay-spectrum technique as the initial approach taken by HERA.  Finally, we provide a short discussion on calibration in the concept of this design.

@@

Since the thermal noise per visibility baseline may be expressed as
\begin{equation}
V_N = \left(\frac{2k_B}{\lambda^2}\right)\left(\frac{T_{sys}}{\sqrt{2Bt}}\right)\Omega_b B
\label{eq:sensitivity_per_baseline}
\end{equation}
where $t$ is the integration time (see {\em e.g.} \citealt{thompson_et_al2001}),
to determine the sensitivity of an instrument to the power spectrum per baseline we can substitute the thermal noise per baseline for the visibility in Eq. \ref{eq:Pk_baseline}.  Further, since an interferometer typically measures many baselines which may be averaged together to improve the signal-to-noise per $k$-bin, the total sensitivity may be approximated as
\begin{equation}
\Delta^2_N (k,z)\approx \frac{k^3X^2(z)Y(z)}{4\pi^2} \left[\frac{T_{sys}^2(z)\Omega_b(z) }{\tau(z) \mathcal{N}_s(k)}\right]
\label{eq:sensitivity}
\end{equation}
where $\mathcal{N}$ represents the improvement in sensitivity based on the array configuration, which may significantly boost the sensitivity if optimized for this measurement (see \citealt{parsons_et_al2012b}).  Dependencies on redshift and wavenumber have been included.

@@

\begin{figure}[h!]
\centerline{
\includegraphics[height=6cm]{plots/boost.pdf}
\includegraphics[height=6cm]{plots/sensratio.pdf}
\includegraphics[height=4cm]{plots/elementSizes.pdf}
}
\caption{Left: Representative boost factors ($\mathcal{N}$ in Eq. \ref{eq:sensitivity}) normalized to a 19-element hexagonal configuration spaced at 14.6m.  For a given element count, one can get about an order of
magnitude improvement in delay-spectrum sensitivity.  Alternatively, for the same sensitivity one can build roughly 1/3 the antennas.  Right:  Power spectrum sensitivities for various arrays normalized to HERA-19.  Here the element size is included, which is shown as an appropriately sized circle.}
\label{fig:boost}
\end{figure}

@@

\subsection{Frequency and Bandwidth}
\label{sec:freqs}
Figure \ref{fig:cosmos} indicates the frequency range requirement to probe the expected timescale of the epoch of reionization using the 21cm line of hydrogen as the probe.  
These limits are derived from to-date complementary probes of reionization which include measurements of
the optical depth to last scattering in the CMB, QSO spectra, Ly-$\alpha$
absorption in the spectra of quasars and gamma ray bursts and the demographics
of Ly-$\alpha$ emitting galaxies
(Figure~\ref{fig:IonHist}). 
Constraints from these probes are still weak:
Ly-$\alpha$ absorption saturates at very small neutral fractions; galaxy
surveys directly constrain only the bright end of the luminosity function and
depend on an unknown escape fraction of ionizing photons to constrain
reionization; CMB measurements probe an integral quantity subject
to large degeneracies. Even when these observations are
combined into a single 95\% confidence region, the bounds remain weak.
For example, $x_{HI}$ spans almost the entire allowable range of [0,1]
at $z=8$. 21\,cm reionization experiments place much tighter constraints on ionization, with the
red band showing the forecasted 95\% confidence region derived from HERA data,
after marginalizing over astrophysical and cosmological parameters.

From these measurements and models, HERA is required to cover a redshift range of $z=$6 to 13, corresponding to a frequency range of 101 - 203 MHz.  We adopt 100-200 MHz for full performance as the requirement.  Extending the science to include the Dark Ages remains a goal so that efforts are on-going to design a feed to increase the lower limit down to at least 70 MHz without compromising the performance in the above requirement range.  The fall-back position for low frequencies is a serial deployment of a scaled version of the HERA feed or to potentially build additional elements specifically for low frequencies.

@@

\subsection{Delay Response}
\label{sec:delayspec}
Early analysis \citep{elementmemo} indicated that to trace the EOR over a wide range of redshift, internal reflections should be attenuated by about 60 dB from the initial incoming wave by about 60 ns, or a voltage standing wave ratio (VSWR) $<$ 1.002 for frequencies $<$ 17 MHz.  This was a conservative value based on estimates of power in spectrally smooth foreground sources from \cite{santos_et_al2005,2008MNRAS.385.2166A, deoliveira2008, jelic_et_al2008, bernardi_et_al2010} and experience with PAPER dealing with foreground systematics.
This provided the basis for the initial design of the element.    For the contemplated feeds on a primary focus antenna, it was found that a focal length/diameter radio ($f/D$) of approximately 0.32 was optimal.  We may therefore estimate the attenuation as a function of delay and diameter and assign this a ``cost function''.

Figure \ref{fig:costfig} summarizes the adopted cost function.  
The vertical axis shows a very steep delay ``cost'' at 60 ns and the horizontal axis is the cost/performance model, as described in the following section.  
The color coding is the product of the two assigned costs, with red showing a high cost normalized to 350 14m elements (clipped at a ratio of 1.25) and purple being the minimum.
The black lines labelled `1', `2' and `3' are the round-trip travel times for 1, 2 and 3 reflections between the feed and the vertex.
To obtain 60 dB by 60 ns we may interpret those lines in one of two ways:  
(a) at a fixed attenuation/reflection, what diameter is required, and
(b) at a fixed diameter how much attenuation per reflection can be allowed.
For 1 round-trip, for the assumed $f/D$ over the diameters of interest, from Fig. \ref{fig:costfig} all delays are less than 60 ns, so this provides no constraint.   If we want to be able to handle 2 round-trips worth of delay, the diameter should be less than 14.2m, which is near the cost minimum.  If we assume 14.2m, then we can estimate that each reflection must average a return loss better than 15 dB, an aggressive target.  And finally, if we need to allow for 3 round-trips, the antenna diameter would have to be less than 9.4m, where the cost is quickly rising.  To handle the delay requirement therefore, the diameter should be less than 14.2m, or the return loss specification becomes very aggressive.

\begin{figure}[t]
\centerline{
\includegraphics[width=0.6\textwidth]{plots/costfig.pdf} 
}
\caption{\small The background coloring with colorbar shows the relative cost/performance for delay and cost at a fixed performance as a function of element diameter.  See \S\ref{sec:delayspec}, \S\ref{sec:cost} and \cite{elementmemo} for details.  A diameter of 14-m (vertical dashed line) is near the cost minimum and consistent with assumption for the delay specification.
\label{fig:costfig}}
\end{figure}

As a key specification, this has been a focus of study with a series of prototypes and analyses based on modeling, and measurements have been conducted to provide much more rigor than this early specification \citep{ewall-wice_et_al2016-EoXLimits,neben_et_al2016,thyagarajan_et_al2016,patra_et_al0216}.  The simple specification set above more appropriately becomes a series of curves for attenuation and delay at different wavenumbers for realistic foreground models.  Figure \ref{fig:delayspec}, adapted from \cite{thyagarajan_et_al2016}, shows the results at $k$ = 0.10, 0.15, and 0.20 $h$/Mpc for the adopted design.  The analysis shows that
the dish design for should allow for EOR detections for $k>0.10 h/$Mpc.  We also see that the initial simple ``60@60'' specification was a overly conservative (shown as the lightly shaded area bounded by the teal line), however it served as a good initial goal.

@@

As discussed in the previous section, the design of the design principles are three-fold:
\begin{itemize}
\item optimize for the delay-spectrum technique of measuring the EOR power spectrum,
\item minimize costs, and
\item design for a limited lifetime of about five years.
\end{itemize}
The first item primarily means that chromatic effects corresponding to delays appropriate for the measurement described above must be below the expected signal level, which essentially determines the focal length.  The second item constrains the diameter and element count, as well as the focal length over diameter ratio ($f/D$), based on a cost function and maximizing sensitivity per element.  And the third item constrains the construction materials and methods and the operational model.  This led to a fixed transit element with a diameter of 14 meters to strike an optimal balance between sensitivity and systematics, as discussed in the previous section.

@@

\subsubsection{Element Construction}
To minimize cost and aid in construction in a remote location, the construction materials are readily available standard construction materials, making for a very cost-effective design.  The feed is supported from three utility poles using line and hardware primarily from boating activities which can handle the load and external environment.  Given the close-packed design, each pole (except for the perimeter) are shared between three antennas.  This both reduces the number of poles, but also makes for a balanced load.  Note that the perimeter poles can be stayed with guy lines if needed.   The rim is also supported off of these poles, with three smaller posts in between such that the outer perimeter is regular dodecagon shape.  Two of the three intermediate posts are shared with adjacent antennas.

@@

\noindent
{\bf Phase 3:} 
A wideband feed is being designed for the next phase of HERA (see Fig. \ref{fig:futureFeed}). This feed will have wide bandwidth (50-250 MHz) and improved performance compare to the current PAPER dipoles. This wider band will open the window of Cosmic Dawn and the latest stages of the Epoch of Re-ionization to HERA. This feed is based on a modified TEM horn concept where the feed?s beam is optimized for the appropriate illumination of the HERA dish to maximize sensitivity. This is to maximize the effective aperture while minimizing the system temperature (receiver noise, ohmic losses, spill-over, etc.).  This feed will as well replace the current PAPER dipoles and will therefore be mechanically compatible with the current dishes being deployed in South Africa.

@@

Furthermore, as shown in \cite{2015ExA....39..567D}, room temperature ultra low receiver noise (<35K including matching noise) above ~100 MHz can be achieved with COTS transistors if a proper feeding mechanism and matching are designed. Despite sky noise is dominant for most of the EoR band, good matching is always critical in wide band systems, and specifically in the HERA case it is important to ensure good power match as well as noise match in order to reduce the effects of unwanted reflections [2]. We envisage having such a feeding mechanism in the new feed where the first stage LNA will be connected directly to the feeding point of the feed antenna.

@@

\subsection{Analog Signal Path}
\label{sec:analog}
HERA's analog signal path from 70--250 MHz with the full performance EOR band of 100-200 MHz, as shown in Fig. \ref{fig:cosmos} and discussed in \S\ref{sec:freqs}, emphasizes spectral smoothness and robustness.  Although the low-frequency sky has an intrinsically bright signal level, in the desired ``cold patches'' the sky temperature is only about 30 K at the upper frequency end, so relatively low receiver temperatures are desired.  Note that at the low end of the band (70 MHz), the sky temperature is about 1000 K and across the EOR band it ranges from about 100-400 K.
With receiver temperatures of 30--100 K readily achievable with ambient temperature electronics however
HERA's low-noise amplifiers (LNAs) may be inexpensive, passively cooled components integrated into the
antenna feed.

@@

The analog signal path comprises the front-end module (FEM) at the feed, the post-amplfier module (PAM) at field-deployed nodes, and the $\sim 35m$ coaxial cable in between.  Also included are the monitor and control aspects to control and measure its state.

\noindent
{\bf Front-End Module: }
The Front-End Module consists of an LNA as the very first stage in the chain for each polarization.  This component for the existing feed design will be a differential low noise amplifier which consists of two sets of LNAs feeding each dipole arm followed by a balun (or hybrid coupler).  The output from this point is then all single ended and mostly matched to 50? impedance.  Whilst it is desirable to be sky noise dominated in the HERA band, meaning for the first stage, the best achievable noise match must be obtained, it must not come at a too great cost to the power match since reducing the ripple across the band is absolutely essential.  This trade-off will be observed carefully to ensure the FEM is both low noise and well matched.  This stage is followed by carefully matched reflection-less filters and further amplifications of the signal.  In order to ensure very low reflection on the line a large matching attenuator will be used at each end of the cable after the bias-Tee circuit.

@@

Another useful feature of the FEM will be an integrated ?Dicke? switching radiometer similar to the EDGES experiment \cite{2012RaSc...47.0K06R}.  The circuit provides the capability of switching between the output of the antenna, a calibrated noise source and a 50? ambient load.  The antenna measures only a fraction of the sky brightness temperature, determined by the matching between the antenna and receiver.  Three spectrometer measurements (antenna, ambient load and noise source) as well as additional lab measurements of the s-parameters and noise-parameters of the FEM prior to deployment then allow good calibration of the output spectrum. An on-board temperature sensor near the noise source will also aid with the calibration of the receiver chain.  The FEM will be housed in a rugged 5x5x10 cm unit which is water and dust resistant.

\noindent
{\bf Post-Amplifier Module: }
The Front-End Module consists of an LNA as the very first stage in the chain for each polarization.  This component for the existing feed design will be a differential low noise amplifier which consists of two sets of LNAs feeding each dipole arm followed by a balun (or hybrid coupler).  The output from this point is then all single ended and mostly matched to 50? impedance.  Whilst it is desirable to be sky noise dominated in the HERA band, meaning for the first stage, the best achievable noise match must be obtained, it must not come at a too great cost to the power match since reducing the ripple across the band is absolutely essential.  This trade-off will be observed carefully to ensure the FEM is both low noise and well matched.  This stage is followed by carefully matched reflection-less filters and further amplifications of the signal.  In order to ensure very low reflection on the line a large matching attenuator will be used at each end of the cable after the bias-Tee circuit.

Another useful feature of the FEM will be an integrated ?Dicke? switching radiometer similar to the EDGES experiment [1].  The circuit provides the capability of switching between the output of the antenna, a calibrated noise source and a 50? ambient load.  The antenna measures only a fraction of the sky brightness temperature, determined by the matching between the antenna and receiver.  Three spectrometer measurements (antenna, ambient load and noise source) as well as additional lab measurements of the s-parameters and noise-parameters of the FEM prior to deployment then allow good calibration of the output spectrum. An on-board temperature sensor near the noise source will also aid with the calibration of the receiver chain.  The FEM will be housed in a rugged 5x5x10 cm unit which is water and dust resistant.

@@

While some applications may utilise SNAPs in multiple-board chasses, for HERA, SNAP is housed in a field-deployable RFI-tight box.
The node enclosure itself is an RFI-tight housing that can handle the signal paths for up to 12 antennas.
The enclosures are placed throughout the array to minimize the total number constrained by the 35\,m maximum cable length. 
For outrigger antennas the delay specification is not critical, so the cable-length to those may be as long as practical.
In total, including outriggers, the HERA-350 design demands 34 nodes, or an average of just over 10 antennas/node.

>>>>

The past three years have seen deep EOR observations with PAPER, the MWA, 
the Giant Metrewave Radio Telescope (GMRT), and LOFAR.
PAPER and the MWA have produced progressively deeper limits
\citep{dillon_et_al2014,dillon_et_al2015,parsons_etal2014,ali_et_al2015}, with PAPER
yielding the first meaningful constraints on the 21\,cm spin temperature during reionization
(Fig. \ref{fig:limits}). %XXX: This figure doesn't really say much about spin temperature.
The inherent challenge is to simultaneously meet stringent sensitivity requirements 
while suppressing foregrounds $\sim5$ orders of magnitude
brighter than the 21\,cm signal, which has been addressed along a number of approaches.  
To address this challenge,  all aspects of the experimental process must be refined and improved, spanning 
calibration \citep[{\em e.g.},][]{zheng_et_al2014, jacobs_et_al2016, 2016arXiv160300607B}, handling foreground contamination
\citep[{\em e.g.},][]{moore_et_al2013,thyagarajan_et_al2015a,pober_et_al2016},
and the interferometer design itself \citep[{\em e.g.},][]{parsons_et_al2012a,dillon_parsons2016}.

@@

The $\kvpr$ components are directly proportional to the baselines and $\kpar$ are proportional to the Fourier transform of the frequency response \footnote{Note that this equivalency is an approximation, which is very good for the short baselines and small bandwidths generally used here.  For a more complete discussion see \cite{liu_et_al2014a,liu_et_al2014b}}.   
The Fourier transform of a frequency spectrum is a delay spectrum, hence the name ``delay-spectrum'' for the technique that uses this approach.  Note that cosmic evolution limits the largest bandwidth (which determines the smallest $\kpar$) to about 10 MHz -- for larger bandwidths the evolution of the Universe begins to impact the result.  For HERA, wavenumbers are dominated by the bandwidth, not the baseline.

Figure \ref{fig:kperf} shows many of the dependencies on wavenumber by the bandwidth, configuration and cosmology by plotting the perpendicular wavenumber ($\kvpr$; black lines), parallel wavenumber ($\kpar$; blue lines) and total wavenumber ($k$; green lines) as a function of redshift for various bandwidths and baselines.  The redshift range is appropriate for the extended frequency range that is the goal for HERA, however $z$=6-13 is the primary region for EOR studies.
The lines are labelled for the assumed bandwidth and/or baseline, depending on the type of wavenumber shown.  Obviously many values may be used, but here a small sampling of values appropriate for HERA are shown.    
Wavenumbers with an assumed bandwidth ($k$ and $\kpar$) are shown as stepped profiles, with the redshift steps equal to the bandwidth at that redshift.
Note that for 100 MHz, the concept of a power spectrum at a given redshift is no longer appropriate, which is why the line is dashed.
Also in Fig. \ref{fig:kperf}, the lime-green shaded area indicates the initial region in which HERA intends to detect and characterize the EOR using the delay-spectrum technique.

@@

If this $\beta$ factor is determined by the chromaticity of the longest baseline, we see that $\beta=1$ and this line is referred to as the ``horizon line'' ({\em i.e.}, the delay of a source at the horizon).  Systematics will push that line such as to close the EOR ``window'' to the upper left ($\beta > 1$).
% above: should "close" be "move"?
If we can completely control systematics and constrain the effects of the foreground to {\em e.g.} the main beam, the line will move to decrease the size of the wedge ($\beta<1$).
Therefore, a key issue to measuring the EOR power spectrum is to understand and minimize systematics.  For further discussion and an analytical derivation, see \citet{zahn_etal2012}, \citet{vedantham_2012}, and \citet{liu_et_al2014b}. In simulation see \citet{datta_etal2010} and \citet{hazelton_et_al2013}, and for observations
\citet{pober_etal2013b}, \citet{2015arXiv150601026P}, \citet{parsons_etal2014}, and \citet{2015arXiv150206016A}.

Note that the unit of the spatial wavenumber $k$ is length$^{-1}$, in this case the relevant length scale is megaparsecs (Mpc; 1 Mpc = 3.086$\times10^{22}$ m).  In order to normalize for updated measurements of the present-day Hubble parameter $H_0$, it is further normalized by a factor $h$, where $H_0=100h$ (km/s)/Mpc, such that the wavenumbers are expressed as $h$/Mpc.  Also note that at the redshifts of interest $X$ has a value of about 160 Mpc/deg and $Y$ has a value of about 16 Mpc/MHz.

@@

The techniques to measure the power spectrum may be broken down into two principle techniques -- delay-space (\S\ref{sec:delayapproach}) and map-making (\S\ref{sec:mapapproach}) -- plus additional hybrid methods.  These are briefly summarized below, with an emphasis on the delay-spectrum technique as the initial approach taken by HERA.  Finally, we provide a short discussion on calibration in the concept of this design.

@@

Since the thermal noise per visibility baseline may be expressed as
\begin{equation}
V_N = \left(\frac{2k_B}{\lambda^2}\right)\left(\frac{T_{sys}}{\sqrt{2Bt}}\right)\Omega_b B
\label{eq:sensitivity_per_baseline}
\end{equation}
where $t$ is the integration time (see {\em e.g.} \citealt{thompson_et_al2001}),
to determine the sensitivity of an instrument to the power spectrum per baseline we can substitute the thermal noise per baseline for the visibility in Eq. \ref{eq:Pk_baseline}.  Further, since an interferometer typically measures many baselines which may be averaged together to improve the signal-to-noise per $k$-bin, the total sensitivity may be approximated as
\begin{equation}
\Delta^2_N (k,z)\approx \frac{k^3X^2(z)Y(z)}{4\pi^2} \left[\frac{T_{sys}^2(z)\Omega_b(z) }{\tau(z) \mathcal{N}(k)}\right]
\label{eq:sensitivity}
\end{equation}
where $\mathcal{N}$ represents the improvement in sensitivity based on the array configuration, which may significantly boost the sensitivity if optimized for this measurement (see \citealt{parsons_et_al2012b}).  Dependencies on redshift and wavenumber have been included.

@@

\begin{figure}[h!]
\centerline{
\includegraphics[height=6cm]{plots/boost.pdf}
\includegraphics[height=6cm]{plots/sensratio.pdf}
\includegraphics[height=4cm]{plots/elementSizes.pdf}
}
\caption{Left: Representative boost factors ($\mathcal{N}$ in Eq. \ref{eq:sensitivity}) normalized to a 19-element hexagonal configuration spaced at 14.6m.  For a given element count, one can get about an order of
magnitude improvement in delay-spectrum sensitivity.  Alternatively, for the same sensitivity one can build roughly 1/3 the antennas.  Right:  Power spectrum sensitivities for various arrays normalized to HERA-19.  The element sizes are indicated at the far right with appropriately sized circles.}
\label{fig:boost}
\end{figure}

@@

\subsection{Frequency and Bandwidth}
\label{sec:freqs}
Figure \ref{fig:cosmos} indicates the frequency range requirement to probe the expected timescale of the epoch of reionization using the 21cm line of hydrogen as the probe.  
These limits are derived from to-date complementary probes of reionization which include measurements of
the optical depth to last scattering in the CMB, QSO spectra, Ly-$\alpha$
absorption in the spectra of quasars and gamma ray bursts and the demographics
of Ly-$\alpha$ emitting galaxies
(Figure~\ref{fig:IonHist}). 
Constraints from these probes are still weak:
Ly-$\alpha$ absorption saturates at very small neutral fractions; galaxy
surveys directly constrain only the bright end of the luminosity function and
depend on an unknown escape fraction of ionizing photons to constrain
reionization; CMB measurements probe an integral quantity subject
to large degeneracies. Even when these observations are
combined into a single 95\% confidence region, the bounds remain weak.
For example, $x_{HI}$ spans almost the entire allowable range of [0,1]
at $z=8$. Twenty-one centimeter reionization experiments place much tighter constraints on ionization, with the
red band showing the forecasted 95\% confidence region derived from HERA data,
after marginalizing over astrophysical and cosmological parameters.

From these measurements and models, HERA is required to cover a redshift range of $z=6$ to 13, corresponding to a frequency range of 101--203 MHz.  We adopt 100--200 MHz for full performance as the requirement.  Extending the science to include the Dark Ages remains a goal so that efforts are on-going to design a feed to increase the lower limit down to at least 70 MHz without compromising the performance in the above requirement range.  The fall-back position for low frequencies is a serial deployment of a scaled version of the HERA feed or to potentially build additional elements specifically for low frequencies.

@@

\subsection{Delay Response}
\label{sec:delayspec}
Early analysis \citep{elementmemo} indicated that to trace the EOR over a wide range of redshift, internal reflections should be attenuated by about 60 dB from the initial incoming wave by about 60 ns, or a voltage standing wave ratio (VSWR) $<$ 1.002 for frequencies $<$ 17 MHz.  This was a conservative value based on estimates of power in spectrally smooth foreground sources \citep{santos_et_al2005,2008MNRAS.385.2166A, deoliveira2008, jelic_et_al2008, bernardi_et_al2010} and experience with PAPER dealing with foreground systematics.
This provided the basis for the initial design of the element.    For the contemplated feeds on a primary focus antenna, it was found that a focal length/diameter radio ($f/D$) of approximately 0.32 was optimal.  We may therefore estimate the attenuation as a function of delay and diameter and assign this a ``cost function''.

Figure \ref{fig:costfig} summarizes the adopted cost function.  
The vertical axis shows a very steep delay ``cost'' at 60 ns and the horizontal axis is the cost/performance model, as described in the following section.
% should this just be "horizontal axis is the element diameter"?
The color coding is the product of the two assigned costs, with red showing a high cost normalized to 350 14m elements (clipped at a ratio of 1.25) and purple being the minimum.
The black lines labelled `1', `2' and `3' are the round-trip travel times for 1, 2 and 3 reflections between the feed and the vertex.
To obtain 60 dB by 60 ns we may interpret those lines in one of two ways:  
(a) at a fixed attenuation/reflection, what diameter is required, and
(b) at a fixed diameter how much attenuation per reflection can be allowed.
For 1 round-trip, for the assumed $f/D$ over the diameters of interest, from Fig. \ref{fig:costfig} all delays are less than 60 ns, so this provides no constraint.   If we want to be able to handle 2 round-trips worth of delay, the diameter should be less than 14.2m, which is near the cost minimum.  If we assume 14.2m, then we can estimate that each reflection must average a return loss better than 15 dB, an aggressive target.  And finally, if we need to allow for 3 round-trips, the antenna diameter would have to be less than 9.4m, where the cost is quickly rising.  To handle the delay requirement therefore, the diameter should be less than 14.2m, or the return loss specification becomes very aggressive.

\begin{figure}[t]
\centerline{
\includegraphics[width=0.6\textwidth]{plots/costfig.pdf} 
}
\caption{\small The background coloring with colorbar shows the relative cost/performance for delay and cost at a fixed performance as a function of element diameter.  See \S\ref{sec:delayspec}, \S\ref{sec:cost} and \cite{elementmemo} for details.  A diameter of 14-m (vertical dashed line) is near the cost minimum and consistent with assumptions for the delay specification.
\label{fig:costfig}}
\end{figure}

As a key specification, this has been a focus of study with a series of prototypes and analyses based on modeling, and measurements have been conducted to provide much more rigor than this early specification \citep{ewall-wice_et_al2016-EoXLimits,neben_et_al2016,thyagarajan_et_al2016,patra_et_al0216}.  The simple specification set above more appropriately becomes a series of curves for attenuation and delay at different wavenumbers for realistic foreground models.  Figure \ref{fig:delayspec}, adapted from \cite{thyagarajan_et_al2016}, shows the results at $k$ = 0.10, 0.15, and 0.20 $h$/Mpc for the adopted design.  The analysis shows that
the dish design for should allow for EOR detections for $k>0.10 h/$Mpc.  We also see that the initial simple ``60@60'' specification was overly conservative (shown as the lightly shaded area bounded by the teal line), however it served as a good initial goal.

@@

As discussed in the previous section, the goals of the design principles are three-fold:
\begin{itemize}
\item optimize for the delay-spectrum technique of measuring the EOR power spectrum,
\item minimize costs, and
\item design for a limited lifetime of about five years.
\end{itemize}
The first item primarily means that chromatic effects corresponding to delays appropriate for the measurement described above must be below the expected signal level, which essentially determines the focal length.  The second item constrains the diameter and element count, as well as the focal length over diameter ratio ($f/D$), based on a cost function and maximizing sensitivity per element.  And the third item constrains the construction materials and methods and the operational model.  This led to a fixed transit element with a diameter of 14 meters to strike an optimal balance between sensitivity and systematics, as discussed in the previous section.

@@

\subsubsection{Element Construction}
To minimize cost and aid in construction in a remote location, the elements are built from readily available standard construction materials, making for a very cost-effective design.  The feed is supported from three utility poles using line and hardware primarily from boating activities which can handle the load and external environment.  Given the close-packed design, each pole (except for the perimeter) is shared between three antennas.  This both reduces the number of poles, but also makes for a balanced load.  Note that the perimeter poles can be stayed with guy lines if needed.   The dish rim is also supported off of these poles, with three smaller posts in between such that the outer perimeter is regular dodecagon shape.  Two of the three intermediate posts are shared with adjacent antennas.

@@

\noindent
{\bf Phase 3:} 
A wideband feed is being designed for the next phase of HERA (see Fig. \ref{fig:futureFeed}). This feed will have wider bandwidth (50-250 MHz) and improved performance compare to the current PAPER dipoles. This wider band will open the window of Cosmic Dawn and the latest stages of the Epoch of Re-ionization to HERA. This feed is based on a modified TEM horn concept where the feed's beam is optimized for the appropriate illumination of the HERA dish to maximize sensitivity. This is to maximize the effective aperture while minimizing the system temperature (receiver noise, ohmic losses, spill-over, etc.).  This feed will as well replace the current PAPER dipoles and will therefore be mechanically compatible with the current dishes being deployed in South Africa.

@@

Furthermore, as shown in \cite{2015ExA....39..567D}, room temperature ultra low receiver noise ($<$35K including matching noise) above ~100 MHz can be achieved with COTS transistors if a proper feeding mechanism and matching are designed. Despite the fact that sky noise is dominant for most of the EoR band, good matching is always critical in wide band systems, and specifically in the HERA case it is important to ensure good power match as well as noise match in order to reduce the effects of unwanted reflections [2].
% this "[2]" is probably a citation that has not been properly propagated??
We envisage having such a feeding mechanism in the new feed where the first stage LNA will be connected directly to the feeding point of the feed antenna.

@@

\subsection{Analog Signal Path}
\label{sec:analog}
HERA's analog signal path from 70--250 MHz with the full performance EOR band of 100--200 MHz, as shown in Fig. \ref{fig:cosmos} and discussed in \S\ref{sec:freqs}, emphasizes spectral smoothness and robustness.  Although the low-frequency sky has an intrinsically bright signal level, in the desired ``cold patches'' the sky temperature is only about 30 K at the upper frequency end, so relatively low receiver temperatures are desired.  Note that at the low end of the band (70 MHz), the sky temperature is about 1000 K and across the EOR band it ranges from about 100--400 K.
With receiver temperatures of 30--100 K readily achievable with ambient temperature electronics however
HERA's low-noise amplifiers (LNAs) may be inexpensive, passively cooled components integrated into the
antenna feed.

@@

The analog signal path comprises the front-end module (FEM) at the feed, the post-amplifier module (PAM) at field-deployed nodes, and the $\sim 35$ m coaxial cable in between.  Also included are the monitor and control aspects to control and measure its state.

\noindent
{\bf Front-End Module: }
The Front-End Module consists of an LNA as the very first stage in the chain for each polarization.  This component for the existing feed design will be a differential low noise amplifier which consists of two sets of LNAs feeding each dipole arm followed by a balun (or hybrid coupler).  The output from this point is then all single ended and mostly matched to 50$\Omega$ impedance.  Whilst it is desirable to be sky noise dominated in the HERA band, meaning for the first stage, the best achievable noise match must be obtained, it must not come at a too great cost to the power match since reducing the ripple across the band is absolutely essential.  This trade-off will be observed carefully to ensure the FEM is both low noise and well matched.  This stage is followed by carefully matched reflection-less filters and further amplifications of the signal.  In order to ensure very low reflection on the line a large matching attenuator will be used at each end of the cable after the bias-Tee circuit.

@@

Another useful feature of the FEM will be an integrated ``Dicke'' switching radiometer similar to the EDGES experiment \citep{2012RaSc...47.0K06R}.  The circuit provides the capability of switching between the output of the antenna, a calibrated noise source and a 50$\Omega$ ambient load.  The antenna measures only a fraction of the sky brightness temperature, determined by the matching between the antenna and receiver.  Three spectrometer measurements (antenna, ambient load and noise source) as well as additional lab measurements of the s-parameters and noise-parameters of the FEM prior to deployment then allow good calibration of the output spectrum. An on-board temperature sensor near the noise source will also aid with the calibration of the receiver chain.  The FEM will be housed in a rugged 5$\times$5$\times$10 cm unit which is water and dust resistant.

\noindent
{\bf Post-Amplifier Module: }
TODO.
% The text here was a duplicate of the FEM subsection.

@@

While some applications may utilise SNAPs in multiple-board chassis, for HERA, SNAP is housed in a field-deployable RFI-tight box.
The node enclosure itself is an RFI-tight housing that can handle the signal paths for up to 12 antennas.
The enclosures are placed throughout the array to minimize the total number constrained by the 35\,m maximum cable length. 
For outrigger antennas the delay specification is not critical, so the cable-length to those may be as long as practical.
In total, including outriggers, the HERA-350 design demands 34 nodes, or an average of just over 10 antennas/node.
